---
title: "Lab 9 Intro"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("jpeg")
```

## Seating
1. If you already have a group, it will be helpful to start sitting next to each other.
2. If you don't have a group, hang tight we are working on it.
3. Please setup your computer as normal for today.

## Warm-up
1. In the data folder there is an epidemiology data set on an outbreak of malaria.

2. `rdt` refers to rapid diagnostic test and they are identified here by age group.

3. Make the data tidy and store them as a new object.
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(here)
library(dplyr)
``` 
```{r}
malaria <- read.csv(here("data","malaria.csv"))
malaria
```
```{r}
malaria2<-malaria%>%
  separate(data_date, into=c("year","month","day"),sep ="-")%>%
  separate(submitted_date, into=c("year","month","day"),sep ="-")
malaria2
```

```{r}
malaria_long<-malaria%>%
  pivot_longer(cols=starts_with("malaria_rdt"),names_to="age_class",values_to="cases")%>%
  select(newid,data_date,submitted_date,location_name,Province,District,age_class,cases)
malaria_long
#select can reorder the columns by the order the columns are listed
```

4. Which district had the highest *total* number of cases on July 30, 2020?
```{r}
malaria2%>%
  filter(year=="2020", month=="07",day=="30")%>%
  group_by(District)%>%
  summarise_at(vars(malaria_tot),list(cases_per_district=sum),na.rm=T)%>%
  arrange(desc(cases_per_district))
#Dingo
#my code's sum do not match ledford's
```
```{r}
malaria_long%>%
  filter(data_date=="2020-07-30")%>%
  group_by(District)%>%
  summarize(tot_cases=sum(cases,na.rm=T))%>%
  arrange(-tot_cases)
```

