---
title: "Lab 7 Intro"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("jpeg")
```

## Setup
1. Login to the lab computer (please don't use your personal computer).  
2. Navigate to github.com and login.   
2. Use GitHub desktop to clone your repository to the desktop.   
5. Copy the class repository to the desktop (https://github.com/jmledford3115/datascibiol).  
6. Copy the files for today's lab from the class repository and paste them into **your** repository.  
7. Open today's lab in RStudio.  

## Review from last time
### *With a partner, discuss the following questions*
1. How did the midterm go? Did you learn anything while doing the exam?
  Felt ok, learned slice_min and slice_max
2. Where did you get stuck?
  mutating the conservation_status column based on the pop size, I got an error saying I couldn't mutate with characters
3. What do we need to practice?
  More practice mutating and filtering/sorting data

### Warm-up
1. Please install the package `naniar`.
```{r}
#install.packages("naniar")
library("naniar")
```

